---
title: "Chapter 3 Practices"
output: html_notebook
---
```{r message=FALSE}
library(rethinking)
library(tidyverse)
```


```{r}
p_grid <- seq(from=0, to=1, length.out=1000)
prior <- rep(1,1000)
likelihood <- dbinom(6, size=9, prob=p_grid)
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample(p_grid, prob=posterior, size=1e4, replace=TRUE)
```


# 3E1
```{r}
sum(samples < 0.2) / length(samples)
```

# 3E2
```{r}
sum(samples > 0.8) / length(samples)
```

# 3E3
```{r}
sum(samples > 0.2 & samples < 0.8) / length(samples)
```

# 3E4
```{r}
quantile(samples, 0.2)
```

# 3E5
```{r}
quantile(samples, 0.8)
```

# 3E6 
```{r}
HPDI(samples, prob=0.66)
```

```{r}
dens(samples)
abline(v=0.5085)
abline(v=0.7738)
```

# 3E7
```{r}
(1-0.66)/2
quantile(samples, c(0.17, 1-0.17))
```




# 3M1
```{r}

#3M1
grid <- seq(0,1, length.out=1000)
p <- rep(1,1000)
W <- dbinom(8,15,prob=grid)
posterior_raw <- p * W
posterior <- posterior_raw / sum(posterior_raw)
```

# 3M2
```{r}
set.seed(123)
samples <- sample(grid, 1e4, replace=TRUE, prob=posterior)
HPDI(samples, prob=0.9)

```

# 3M3
```{r}
dum_data <- rbinom(1e4,15,prob=samples)
prob <- length(dum_data[dum_data== 8]) / length(dum_data)
cat("The probability of getting 8 out of 15 is:" , prob)
```

# 3M4
```{r}

# so, according to jffist's solution, there are two ways to carry this out. 

# here is approach 1, but I don't understand this tho
likelihood6of9 <- dbinom(6,9,grid)
prior6of9 <- posterior
posterior6of9 <- likelihood6of9 * prior6of9
cat("The probability of getting 6 out of 9 is:", sum(posterior6of9))




```

```{r}
# Here is approach 2, which uses generative approach
set.seed(22)
dummy <- rbinom(n=1e4, size=9, prob=samples)
sum(dummy == 6) / length(dummy)
```

# 3M5

## 3M5.1
```{r}

prior <- ifelse(grid < 0.5,0, 2)
likelihood <- dbinom(8,15,prob=grid)
posterior_raw <- prior * likelihood
posterior <- posterior_raw / sum(posterior_raw)
```


## 3M5.2
```{r}
samples <- sample(grid, 1e4, replace=TRUE, prob=posterior)
HPDI(samples, prob=0.9)
```

## 3M5.3
```{r}
dummy <- rbinom(1e4,15,prob=samples)
sum(dummy==8) / length(dummy)
```

# 3M6 
This problem is done in the homework of week 1

***

```{r}
data(homeworkch3)
sum(birth1) + sum(birth2)
```

# 3H1

```{r}
birth_all<- c(birth1, birth2)
grid <- seq(0,1, length.out=1000)
prior <- rep(1,1000)
likelihood <- dbinom(x=sum(birth_all), size=length(birth_all), prob=grid)
posterior_raw <- prior * likelihood
posterior <- posterior_raw / sum(posterior_raw)
grid[which.max(posterior)]
```


# 3H2
```{r}
samples <- sample(grid, 1e4, replace=TRUE, prob=posterior)
HPDI(samples, prob=0.5)
HPDI(samples, prob=0.89)
HPDI(samples, prob=0.97)
```


# 3H3

```{r}
dummy <- rbinom(1e4, 200, prob=samples)
dens(dummy)
mean(dummy)
median(dummy)
```

The max frequency in the simulated data is 111, which shows that this model fits the data pretty well

# 3H4

Is there any quantatitive way of comparing these two samples? I hope there is. 

But in term of frequencies, the first set of data:
```{r}
sum(birth1) / length(birth1)
```

Now calculate the frequencies in the sample
```{r}
mean(dummy/ 200)
```

which is slightly higher than the first set of data




# 3H5
```{r}
girl_cnt <- sum(birth1==0)
dummy <- rbinom(1e4, girl_cnt, samples)
mean(dummy)
sum(birth2[which(birth1==0)])
```

So I guess people selected by either aborsion or doing something else... I don't know and I don't want to think what kind of egregous deeds that people can do in this case.
